Distribution-oblivious Programming with Dish

### 2019-11-20 pldi/draft.tex
Update overview

@@

>>>>

As outlined earlier, the key insight behind \sys is that the shell already describes parallel computations over streams.
Intuitively
 (i) the parallel-composition operator exposes task parallelism,
 (ii) commands such as \ttt{xargs} and \ttt{tee} expose data parallelism,
 (iii) the pipe operator enables both task and data parallelism, and
 (iv)  other composition operators can be viewed as barriers.
To leverage these insights and correctly distribute a pipeline such as $p_1$, \sys must solve several challenges.

Understanding the primitives of the shell is not enough;
  one has to understand the characteristics of commands used to compose larger programs.
This challenge has two parts
One, characterize the commands already in the shell---that is, built-ins .
This study identifies a small set of well-understood distributability classes and is already useful enough to allow the composition of practical pipelines.
The second part is to use these classes to 
Using this language -> extensions
A command developer can only annotate the default behavior;
  if a new flag is added a few years later, and only if that flag alters the command's class, only then a new annotation is needed.

Given a set of , \sys's has to analyze the given pipeline and identify candidate subexpressions for distribution.
\sys addresses this challenge with a dataflow analysis that identifies longest subexpressions.
This analysis is not trivial, as many of the shell's dynamic features have to be taken into consideration---examples include shell expansion, environment variables, file input sizes, \etc

Given a distributed pipeline with runtime instrumentation---two challenges.
First, wrapping 
  just-in-time wrappers for all the, generated by the distributability specification outlined earlier.
The reasoning behind thin wrappers is that they should not interfere with legacy functionality---commands have many flags accumulated over long periods of time.
Second, fill-in planing holes left by the rewriting engine.
This is achieved with \emph{planner activators}, commands yield (passing their arguments) to the scheduler.
From the point of view of the shell, activators are higher-order commands a la \ttt{xargs} and \ttt{time} that take other commands as arguments---a technique that achieves the latest possible binding.

The next few sections , and discuss other challenges such as the synchronization of environment variables, distributed file operation, \etc.

\section{Distributability Classes}

@@

Here is the type for \ttt{bwa}, a command that performs a Burrows-Wheeler transform over genomic data:
% ( bwa ) :: ( Pure )  => { In \/ File* } -> { Out /\ Err }    -- default case
%    | -h :: ( Sless ) => { Out }                              -- if needed
\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
(bwa) :: (Pure) ($$\Rightarrow$$){In($$\lor$$)File*}($$\rightarrow$$){Out($$\land$$)Err}
  |-h :: (Sless)($$\Rightarrow$$){Out}                              
\end{lstlisting}

It states that the command defaults to the Pure class.
It's operation can be thought as a transformation from input streams to output streams.
It operates either on stdin or, if this doesn't exist, one or more files specified as arguments;
% TODO: check actual man page for how to specify files
  it writes to the output and error stream rather than a file.
The \ttt{-h} flag moves \ttt{bwa} into the Stateless class;
  its output is a constant function.
